---
title: "DATA 607 Week 9 Assignment - Web APIs"
author: "Stephen Haslett"
date: "10/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Overview
The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in Python to read in the JSON data, and transform it into a pandas DataFrame.

For this assignment, I chose to connect to the New York Times' Article Search API. I queried the API for articles relating to vaping deaths, and limited the search results to articles relating to the keyword "Death".

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(jsonlite)
library(tidyr)
library(dplyr)
library(stringr)
library(kableExtra)
```

###Connect to the NYT Articles API and retrieve articles relating to vaping deaths.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
api_key <- "&api-key=sm2AdtUSMtxAW5ZDlRLYAAGyYMrYGTnn"
search_query <- "?q=vaping%20deaths"

# Define filters to limit the search results.
filters <- "&fq=document_type(article)&fq=type_of_material(News)&fq=keyword(Deaths (Fatalities))"

sort_by <- "&sort=relevance"
start <- "&begin_date=20190101"
end <- "&end_date=20191026"
article_endpoint <- "https://api.nytimes.com/svc/search/v2/articlesearch.json"

vaping_articles <- fromJSON(paste0(article_endpoint, search_query, filters, start, end, sort_by, api_key))
results <- vaping_articles$response$docs
```


###Output the results in a data frame.
```{r, eval=TRUE, message=FALSE, warning=FALSE}
vaping_df <- data.frame('Date' = format(as.Date(results$pub_date), "%B %d, %Y"),
                        'Headline' = results$headline$main,
                        'Abstract' = results$abstract,
                        'URL' = results$web_url)

vaping_df %>% kable() %>% kable_styling() %>% scroll_box(width = "910px", height = "400px")
```