---
title: "Week 12 Assignment - NoSQL Migration"
author: "Stephen Haslett"
date: "11/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RMySQL)
library(mongolite)
library(dplyr)
library(stringr)
```

##Assignment Description
For this assignment, you should take information from a relational database and migrate it to a NoSQL database of your own choosing.

For the relational database, you might use the flights database, the tb database, the "data skills" database your team created for Project 3, or another database of your own choosing or creation.

For the NoSQL database, you may use MongoDB (which we introduced in week 7), Neo4j, or another NoSQL database of your choosing.

Your migration process needs to be reproducible. R code is encouraged, but not required. You should also briefly describe the advantages and disadvantages of storing the data in a relational database vs. your NoSQL database.

###Step 1 - Import the MYSQL flights database data into R.


#### Connect to the MYSQL Flights database.
```{r, warning=FALSE, message=FALSE}
# Open the database connection and list the database tables. 
connection <- dbConnect(RMySQL::MySQL(), user = 'root', password = '', dbname = 'flights', host = 'localhost')
tables <- dbListTables(connection)

tables
```


#### Import the MYSQL table data into R, and store it in dataframes.

**_For this assignment, I am only interested in the "airlines", "airports", and "flights" tables, so I will ignore the "planes" and "weather" tables_.**

```{r, eval=FALSE, warning=FALSE, message=FALSE}
airlines <- dbReadTable(connection, 'airlines')
airports <- dbReadTable(connection, 'airports')
flights <- dbReadTable(connection, 'flights')

# Close the database connection.
dbDisconnect(connection)
```


#### Confirm that the database table data is now available in R.
```{r, warning=FALSE, message=FALSE}

head(airlines)
```


#### Remove the carriage returns ("\r") from the "name" column of the airlines dataframe so that the data is cleaner.
```{r, warning=FALSE, message=FALSE}
airlines$name <- str_replace(airlines$name, '\\r', '')
colnames(airlines)
head(airlines)
```


```{r, warning=FALSE, message=FALSE}
head(airports)
```


```{r, warning=FALSE, message=FALSE}
head(flights)
```


### Step 2 - Import the Data into MongoDB.


#### Prepare the flights data for insertion into the MongoDB database by merging the dataframes together.
```{r, warning=FALSE, message=FALSE}
# Merge the 3 dataframes (airlines, airports, and flights) together with the help of the dplyr package.
# We create a new object (flights_data) to contain the merged data. This object will then be imported into the MongoDB database.

# Add the flights dataframe to the flights_data object, and merge in the airlines dataframe.
flights_data <- flights %>%
  merge(airlines) %>%
  # To prevent conflicts with the "airports" table name column, we rename the "airlines" name column to "airline".
  rename(airline = name) %>%
  rowwise() %>%
  ungroup()

# Now add the airports dataframe, and seperate the data into origin and destination airports.
flights_data <- airports %>%
  rename_(.dots = setNames(names(.), paste0('origin_', names(.)))) %>%
  rename(origin = origin_faa) %>%
  right_join(flights_data)

flights_data <- airports %>%
  rename_(.dots = setNames(names(.), paste0('dest_', names(.)))) %>%
  rename(dest = dest_faa) %>%
  right_join(flights_data)
```


#### Open a connection to the MongoDB database.
```{r, warning=FALSE, message=FALSE}
mongo_database <- mongo(collection = "flights_data", db = "flights_mongo")
```

#### Insert the flights data into the MongoDB database.
```{r, warning=FALSE, message=FALSE}
mongo_database$insert(flights_data)
```


####Confirm that the data has been successfully imported into MongoDB by running a query for flights flying out of JFK.
```{r, warning=FALSE, message=FALSE}
results <- mongo_database$find('{"origin": "JFK"}') %>%
  group_by(airline, origin_name, dest_name) %>%
  summarise(total_flights = n()) %>%
  arrange(airline)

results
```

